\chapter{Discussion}
\section{Tau}
\input{Content/tau_discussion}
\section{Ase1}
\input{Content/ase1_discussion}

\section{The patterning of cytoskeletal polymers}
Can we draw any insights which go beyond the pecuilarities of each of the two MAPs I was focusing on? Overall, the findings associated with both the Ase1-related experiments and the tau-related experiments both lend further support to the picture of a plastic microtubule lattice. 

This multimerization, particularly when enhanced by Ase1 herding at depolymerizing MT ends, could introduce additional complexity to Ase1-mediated MT dynamics regulation. The possibility of Ase1 molecules acting cooperatively to promote rescues for antiparallel MTs specifically is intriguing and would offer the cell a lever for modulating the rescue-promoting effect of MT crosslinking.


- If tubulin bonds were strong laterally, then the cell would have to break up two strong bonds in order to initiate dissassembly -> Maybe would be too complicated, which is why such stabilization does not really take place a lot?
- On the importance of interplay and cooperation?

. Valiron O, Arnal I, Caudron N, Job D: GDP-tubulin incorporation
into growing microtubules modulates polymer stability. J Biol
Chem 2010, 285:17507-17513.
13. Bowne-Anderson H, Zanic M, Kauer M, Howard J: Microtubule
dynamic instability: a new model with coupled GTP hydrolysis
and multistep catastrophe. Bioessays 2013, 35:452-461.
14. Piedra F-A, Kim T, Garza ES, Geyer EA, Burns A, Ye X, Rice LM:
GDP-to-GTP exchange on the microtubule end can contribute
to the frequency of catastrophe. Mol Biol Cell 2016, 27:3515-
3525.