\chapter*{Abstract}
The cytoskeleton is a complex network of interlinking protein filaments which fulfills a diverse range of functions for the cell. One of the key type of protein filaments are microtubules, which themselves are present in a diverse set of cellular contexts. In axons, long, stable microtubule arrays within the microtubule shaft form highways for molecular motors, enabling them to efficiently transport cargo across the cell. \textit{In vitro} experiments, i.e., experiments where cellular components of interest are investigated outside of their cellular context, can help shed light on potential microscopic mechanisms underlying and giving rise to observed macroscopic phenomena. The results of the \textit{in vitro} experiments conducted for this thesis contribute to our understanding how these arrays could potentially be stabilized and protected against microtubule-severing enzymes while other microtubule regions within the neuron can remain dynamic. In particular, the results presented here show how the microtubule-associated protein Tau, which preferentially locates to axonal shafts, cooperatively binds to microtubules, forming islands which are highly effective at protecting microtubules against severing. We also show that these Tau islands regulate the activity of the molecular motor kinesin-8, and that this motor in turn is also capable of disassembling Tau islands. In the second \textit{in vitro} case study presented in this thesis, we show that the microtubule-crosslinking protein Ase1 can selectively stabilize antiparallel microtubule arrays, as they are found during mitosis, against depolymerization from the microtubule ends. This case study also shows that Ase1 is being herded by depolymerizing microtubule ends, and that this phenomenon likely is related to the propensity of Ase1 to oppose microtubule depolymerization. Overall, this the results in this thesis contribute to our understanding of how distinct interaction patterns between microtubules and microtubule-associated proteins can give rise to macroscopic structures within the cell.

\chapter*{Abstrakt}
Cytoskelet je složitá síť vzájemně propojených proteinových filament, která plní v buňce celou řadu funkcí. Jedním z klíčových typů proteinových filament jsou mikrotubuly. Mikrotubuly v axonech tvoří dlouhé, stabilní struktury podél nichž se pohybují molekulární motory, které umožňují transport buněčného nákladu, jako jsou například organely, napříč buňkou. Experimenty in vitro, tj. experimenty, při nichž jsou zkoumány buněčné složky mimo jejich buněčný kontext, mohou pomoci objasnit potenciální molekulární mechanismy, které jsou základem a příčinou pozorovaných makroskopických jevů. Výsledky první in vitro studie provedené v rámci této práce přispívají k pochopení toho, jak mohou být mikrotubuly stabilizovány a chráněny v určité části buňky, zatímco mikrotubuly v jiné oblasti mohou zůstat dynamické. Předkládané výsledky ukazují, jak se mikrotubulární protein Tau, který se lokalizuje zejména na axonální mikrotubuly, kooperativně váže na tyto mikrotubuly a vytváří ostrůvky, které ochraňují a stabilizují mikrotubuly. Ukázali jsme také, že tyto Tau ostrůvky regulují aktivitu molekulárního motoru kinesinu-8 a že tento motor je také schopen tyto ostrůvky rozkládat. Ve druhé in vitro studii, kterou předkládáme v této práci, ukazujeme, že mikrotubulární spojovací protein Ase1 selektivně stabilizuje antiparalelní mikrotubulární struktury, které jsou například v centru mitotického vřeténka. Naše výsledky ukazují, že molekuly Ase1 jsou během depolymerace mikrotubulu shromážděné pomocí depolymerujícího konce mikrotubulu, což má za následek zpomalení depolymerace. V souhrnu, výsledky prezentované v této práci ukazují jak interakce mezi mikrotubuly a mikrotubulárními proteiny mohou ovlivňovat mikrotubulární dynamiku a tím buněčné pochody.
